### 2008-08-04 doc/a-rex/arex_tech_doc.tex
Started removing gridftp and grid-manager from a-rex documentation

%% LyX 1.4.5 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[twoside,english]{article}
\usepackage{pslatex}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{geometry}
\geometry{verbose,a4paper,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2cm}
\usepackage{array}
\usepackage{amsmath}
\usepackage{setspace}
\onehalfspacing
\usepackage{amssymb}
\IfFileExists{url.sty}{\usepackage{url}}
                      {\newcommand{\url}{\texttt}}

@@

A-Rex is one of ARC middleware componets implementing functions of
so called \emph{Computing Element} (CE). It's made of few components
- A-Rex Web Service (WS) Interface and Grid Manager (GM). Alternative
and currently only stable interface to A-Rex is GridFTP Server (GFS).

The aim of the Grid Manager (GM) is to take care of job pre- and post-processing.
It provides functionality to stage-in files containing input data
and program modules from wide range of sources and transfer or store
output results.

Both WS Interface and GridFTP Interface components provide a way to
submit and control computational tasks (jobs) to be executed by the
GM and undelying Local Resource Management System. While WS interface
is currently being developed the GridFTP is considered to be stable
and prefered one.

@@

Each job gets a directory on the CE called the \emph{session directory}
(SD). Input files are gathered in the SD. The job is supposed to produce
new data files  in the  SD as well. GM does not guarantee the
availability of any other places accessible by the job other than
SD (unless such a place is part of a requested Runtime Environement).
The SD is also the only place which is controlled by the GM. It is
is accessible by the user from outside through HTTPS and GridFTP protocols.
Any file created outside the SD is not controlled by the GM. Any exchange
of data between client and GM (including also program modules) is
done through GridFTP \cite{gridftp} or HTTPS protocols depending
on the selected interface. A URL for accessing input/output files
is obtained through the Information System or the WS Information Interface
of A-Rex.

Each job gets an identifier (\textit{jobid}). This is a handle which
identifies the job in the GM and the Information System \cite{is}.

  Jobs are initiated and controlled through GFS or WS. Complete
job descriptions   (JD)   are passed to the GM either
through GFS in RSL \cite{rsl} or through WS in JSDL \cite{jsdl}
coded description     . Input data files
and job executables are transferred separately through the same interface,
as described in the next section.

@@

The main task of the GM is to take care of processing input and output
data (files) of the job. Input files are gathered in the SD. There
are two ways to put a file into the SD:

\begin{itemize}
\item Download is initiated by the GM --  This is the case for
files     defined in the JD (with name and source).
      The GM alone is responsible to make
sure that all required files will be available in the SD.\\
The supported protocols for sources  at the moment are: GridFTP,
FTP, HTTP, HTTPS (HTTP over SSLv3) and HTTPg (HTTP over GSI). Also
some nonstandard sources are supported. Those are described below.
\item Upload is initiated by the user directly or through the User Interface
(UI). Because the SD becomes available immediately at the time of
submission of JD, UI can (and should) use that to upload data files
which are not otherwise accessible by the GM.  Examples of such
files  are the main executable of the job,   the
job's input files, etc. These files can (and should)
also be specified in the JD.
\end{itemize}
\textbf{There is no} other reliable way for a job to obtain input
data on the CE belonging to NorduGrid. Access to AFS, NFS, FTP, HTTP
and any other remote data transport during execution of a job is
not guaranteed (at least not by GM).

Jobs should store output files in thir SD. Like input files,
output files  belong into two groups:

\begin{itemize}
\item Files which are supposed to be moved to a \emph{Storage Element} (SE)
and optionally registered in some \emph{Indexing Service} like the
Globus \emph{Replica Location Service } (RLS)  --
The GM takes care of these files. They have to be specified in the
JD. If the job fails during any stage of processing, no attept is
made to transfer those files to their final destination, unless
the  option \emph{preserve=yes} is specified in their URLs.
\item Files which are supposed to be fetched by the user -- The user has
to use a tool like the UI to obtain these files. They \textbf{must}
also be specified in the JD.
\end{itemize}

@@

From the point of view of the GM a job passes through various states.
Picture \ref{job states diagram} presents a diagram of the possible
states of a job.%
\begin{figure}[tbh]
\begin{centering}
\includegraphics{pic1}
\par\end{centering}

\caption{\label{job states diagram}Job states}
\end{figure}
A user can examine the state of a job by querying the NorduGrid Information
System (IS) using the UI or any other suitable tool. Please remember
that the IS can manipulate  state names to make them more user
friendly and to combine them with states introduced by other parts
of the whole setup. Another way is to access virtual informational
files through GridFTP interface or to use query method of WS interface.

@@

Below is the description of all actions taken by the GM at every state:

\begin{itemize}
\item \textbf{Accepted} - In this state the job has been submitted to
a CE but is not processed yet. The GM will analyze the JD and move
to the next stage. If the JD can not be processed the job will be
canceled and moved to the state \textbf{Finishing}.
\item \textbf{Preparing} - The input data is being gathered in the SD (stage-in).
The GM is downloading the files specified in the JD and is waiting
for files which are supposed to be uploaded by the UI. If
all files are successfully gathered the job moves to the next state.
If \textbf{any} file can't be downloaded or it takes the UI too long
to upload a file,  the job moves to \textbf{Finishing} state. It
is possible to put a limit on the number of simultaneous \textbf{Preparing}
jobs. If this limit is exceeded, jobs ready to enter the \textbf{Preparing}
state      will stay in the \textbf{Accepted}
state, but prefixed with the PENDING: mark. Exceptions are jobs which
have no files to be downloaded. These are processed out of limits.
\item \textbf{Submitting} -      The job is being
passed for execution to the \emph{Local Resource Management System}
(LRMS). At the moment PBS is supported best but correspoding backends
for many other LRMSs are provided with the default installation. 
      If the local job submission
is successful the job moves to the \textbf{InLRMS} state. Otherwise
it moves to \textbf{Finishing}. It is possible to limit the agregate
number of jobs in \textbf{Submitting} and  \textbf{InLRMS}
states.
\item \textbf{InLRMS} - The job is queued or being executed in the LRMS.
The GM takes no actions except waiting until the job finishes.
\item \textbf{Cancelling} - Necessary action to cancel the job in the LRMS
is being taken.
\item \textbf{Finishing} - The output data is being processed (stage-out).
Specified data files are moved to the specified SEs and are optionally
registered at  an Indexing Service. The user can download data files
from the SD by using the UI or  other adequate tool. All the files
not specified as output files are removed from the SD at very beginning
of this state. It is possible to limit number of simultaneous jobs
in this state.
\item \textbf{Finished} - No more processing is performed by the GM. The
user can continue to download data files from the SD. The SD is kept
available for some time (default is 1 week). After that the job
is moved to the state \textbf{Deleted}. The 'deletion' time can be
obtained by querying the NorduGrid Information System  attribute
\texttt{nordugrid-pbs-job-sessiondirerasetime} of a \texttt{nordugrid-pbs-job}
entry. If a job was moved to \textbf{Finished} because of failure,
it may be restarted on request of a client. When restarted, a job
is moved to the state previous to the one in which it failed and
is assigned mark PENDING. This is needed in order to not break the
configuration limits. Exception is a job failed in \textbf{InLRMS}
state and lacking input files specified in JD. Such a job is treated
like failed in \textbf{Preparing} state.
\item \textbf{Deleted} - The job is moved to this state if the user have
not requested job to be cleaned before the SD's lifetime expires.
Only minimal subset of information about such job is kept. The SD
is not available anymore.
\end{itemize}
In  case of a failure special processing is applied to output
files. All specified output files are treated as \textbf{downloadable
by the user}. No files will be moved to their destination SE.

@@

The GM and it's components support the following data transfer protocols
and corresponding URLs: \emph{ftp, gsiftp, http, httpg, https, se,
rc} and \emph{rls.} For more information please see {}``Protocols,
Uniform Resource Locators (URL) and extensions supported in ARC''
document \cite{urls}.

@@

\subsection{Internal Files of the Grid Manager}

For each local UNIX user listed in the GM configuration a \textit{control
directory} exists. In this directory the GM stores information about
jobs belonging to that user. Multiple users can share the same \textit{control
directory}. To make it easier to recover in  case of failure, the
GM stores most information in files rather than in memory. All files
belonging to same the job have names starting with \textbf{job.ID.,}
where ID is the job identifier.

@@

\begin{itemize}
\item \textit{job.ID.description} - contains the RSL description of the
job.
\item \textit{job.ID.local} - information about the job used by the GM.
It consists of lines of format \textit{{}``name = value''} . Not
all of them are always available. The following names are defined:

@@

\item \textit{clientsoftware} - version of software used to submit the job
\item \textit{sessiondir} - the job's SD  
\item \textit{failedstate} - state in which job failed (available only
if it is possible to restart the job)
\item \textit{jobreport} - URL of a user requested \emph{logger service}.
The GM will also send job records to this service in addition to the
default logger service configured in the configuration

\end{itemize}
\end{itemize}
This file is filled partially during job submission and fully when
the job moves from the \textbf{Accepted} to the \textbf{Preparing}
state.

@@

\item \textit{job.ID.proxy} -delegated GSI proxy.
\item \textit{job.ID.proxy.tmp} - temporary GSI proxy with different unx
ownership used by processes run with effective \emph{user id} different
form job owner's \emph{id}.
\end{itemize}
There are other files with names like job.ID.{*} which are created
and used by different parts of the GM. Their presence in the \textit{control
directory} can not be guaranteed and can change depending on changes
in the GM code.

\subsection{GridFTP Interface}

The GridFTP Interface to job management is implemented through GridFTP
server. The GFS is made of layer implementing GridFTP protocol and
plugins for accessing actual data. Local file access in the GFS is
implemented through plugins (shared libraries). There are 3 plugins
provided with the GFS: \textit{fileplugin.so, gaclplugin.so} and \textit{jobplugin.so}
.

The \textit{fileplugin.so} is intended to be uses for plain file access
with the configuration senitive to user subject and is not needed
for job management interface.

The \textit{gaclplugin.so} uses GACL (\url{http://www.gridpp.ac.uk/authz/gacl/})
to control access to local file system instead of Unix user identity.
It's also not needed for job management.

The \textit{jobplugin.so} is using information about jobs being controlled
by GM and provides access to session directories of the jobs owned
by user. It also provides an interface (virtual directory and virtual
operations) to submit, cancel, clean, renew credentials and obtain
information about the job.

@@

A-REX Web Service Interface provides means to submit a description
of a computational job to a computing resource, to stage-in additional
data, to monitor and control processing of jobs, and obtain data produced
during the execution of a job. The WS Interface is built and deployed
inside Hosting Envirnoment Daemon {[}HED] infrastructure\cite{hed}.

@@

The job submission and control interface is based on document produced
by the OGF OGSA Basic Execution Services (BES) Working Group\cite{bes}.
The exchange of SOAP messages is performed over the HTTPS protocol.
The BES interface is represented by two port-types - BES-Management
and BES-Factory. The former is made to control the A-REX service itself
and thus defines operations to start and stop the functionality of
the BES service. A-REX does not implement remote control of service
functionality. Hence the BES-Management port-type is not implemented.
The BES-Factory port-type provides operations to submit new jobs (to
create an activity in terms of BES) and to monitor its state. It also
has an ability to provide information about the service. A-REX fully
implements the functionality of this port-type.

For job descriptions A-REX accepts the Job Submission Description
Language (JSDL)\cite{jsdl} as defined by the OGF Job Submission Description
Language Working Group. Supported elements and extensions are described
below.

@@

A-REX introduces a new operation in addition to those provided by
BES. It does that by defining its own port-type A-REX with the single
operation \emph{ChangeActivityStatus} (see Appendix D). This operation
provides a way to request simple transfers between states of jobs
being processed and corresponding actions.

@@

Since BES allows implementators to extend their initial activity states
with additional sub-states, A-REX defines a set of sub-states of activity
processing in addition to those defined by the BES, as listed in Table
\ref{tab:Job-states-definitions}.

@@

ARC service interfaces optionally offer a sub-interface, called the
Delegation Interface (see Appendix E). This is a common purpose interface
to be used by ARC services which accept delegated credentials from
clients. The Delegation Interface implements two operations: initiatialization
of credentials delegation (DelegateCredentialsInit) and update/renewal
of credentials (UpdateCredentials).

@@

The A-REX implements the Local Information Description Interface (LIDI)
interface being common for all ARC services. This interface is based
on OASIS Web Services Resource Properties specification\cite{wsrf-rp}.
Information about resources and maintained activities/jobs are represented
in a \emph{WS-Resource Properties} informational XML document. The
document type is defined in the A-Rex WSDL as a \emph{ResourceInformationDocumentType}.
It contains the following elements/resources:

@@

\emph{grid-manager {[}-h] {[}-C level] {[}-d level] {[}-c path] {[}-F]
{[}-U uid{[}:gid]] {[}-L path] {[}-P path]}\\
\hspace*{0.5cm}\emph{-h} - short help,\\
\hspace*{0.5cm}\emph{-d} - debug level,\\
\hspace*{0.5cm}\emph{-L} - name log file (overwrites value in configuration
file),\\
\hspace*{0.5cm}\emph{-P} - name for file containing process id (overwrites
value in configuration file),\\
\hspace*{0.5cm}\emph{-U} - user and gropu id to use for running daemon,\\
\hspace*{0.5cm}\emph{-F} - do not make process daemon,\\
\hspace*{0.5cm}\emph{-c} - name od configuration file,\\
\hspace*{0.5cm}\emph{-C} - remove old information before starting:
1- remove finished jobs, 2 - remove active jobs too, 3- also remove
everything that looks like junk.\\

\subsection{\label{subsec: configfile of the gfs}Configuration of the GridFTP
Server}

Default location of the GFS configuration file is \textit{/etc/arc.conf}
or \textit{\$NORDUGRID\_LOCATION/etc/gridftpd.conf}. Format of these
configuration files is similar to that of the GM. It also supports
generic commands described at the beginning of previous section \ref{SubSection: ConfigFile}.
In the new format sections {[}common] and {[}gridftpd] are used. Commands
specific to the GFS are described below.

\begin{itemize}
\item \textbf{\textit{port}}\textit{=number} - specifies TCP/IP port number.
Default is 2811.
\item \textbf{\textit{include}}\textit{=path} - include contents of another
file. Generic commands can't be specified there.
\item \textbf{\textit{encryption}}\textit{=yes|no} - specifies if server
will allow data transfer to be encrypted. Default is yes.
\item \textbf{\textit{pluginpath}}\textit{=path} - specifies the path where
plugin libraries are installed
\item \textbf{\textit{allowunknown}}\textit{=yes|no} - if set to \emph{yes}
clients are not checked agains grid-mapfile. Hence only access rules
specified in this configuration file will be applied.
\item \textbf{\textit{firewall}}\textit{=hostname} - use IP address of the
\textit{hostname} in response to PASV command instead of IP address
of a network interface of computer. You can write IP address directly
instead of \textit{hostname}. This command may be if server is situated
behind NAT.
\item \textbf{\textit{unixgroup}}\textit{=group rule} - define local UNIX
user and optionally UNIX group to which user belonging to specified
authorization \textit{group} is mapped (see Section \ref{sub:Authorization}
for definition of group). Local names are obtained from specified
\textit{rule}. If specified rule could not produce any mapping, next
command is used. Mapping stops at first matched rule. Following rules
are supported:

\begin{itemize}
\item \textbf{\textit{mapfile}} \textit{file} - user's subject is matched
against list of subjects stored inspecified file, one per line followed
by local UNIX name.
\item \textbf{\textit{simplepool}} \textit{directory} - user is assigned
one of local UNIX names stored in a file \textit{directory/pool},
one per line. Used names are stored in other files placed in the same
\textit{directory}. If UNIX name was not used for 10 days, it may
be reassigned to another user.
\item \textbf{\textit{lcmaps}} \textit{library directory database} - call
LCMAPS functions to do mapping. Here \textit{library} is path to shared
library of LCMAPS, either absolute or relative to \textit{directory};
\textit{directory} is path to LCMAPS installation directory, equivalent
of LCMAPS\_DIR variable; \textit{database} is path to LCMAPS database,
equivalent to LCMAPS\_DB\_FILE variable. Each arguments except \textit{library}
is optional and may be either skiped or replaced with '{*}'.
\item \textbf{\textit{mapplugin}} \textit{timeout plugin} {[}\textit{arg1}
{[}\textit{arg2} {[}...]]] - run external \emph{plugin} executable
with specified arguments. Execution of \emph{plugin} may not last
longer than \emph{timeout} seconds. Rule matches if exit code is 0
and there is UNIX name printed on \emph{stdout}. Name may be optionaly
followed by UNIX group separated by ':'. In arguments following substitions
are applied before plugin is started:

\begin{itemize}
\item \%D - subject of users's cerificate,
\item \%P - name of credentials' proxy file.
\end{itemize}
\end{itemize}
\item \textbf{\textit{unixvo}}\textit{=vo rule} - same as \textbf{\textit{unixgroup}}
for users belonging to Virtual Organization (VO) \textit{vo}.
\item \textbf{\textit{unixmap}}\textit{={[}unixname]{[}:unixgroup] rule}
- define local UNIX user and optionally group used to represent connected
client. \textit{rule} is one of those allowed for \textbf{\textit{authorization
groups}} (see Section \ref{sub:Authorization}) and for \textbf{\textit{unixgroup}}/\textbf{\textit{unixvo}}.
In case of mapping rule username is one, provided by rule. Otherwise
specified \textit{unixname:unixgroup} is taken. Both \textit{unixname}
and \textit{unixgroup} may be either omited or set to '{*}' to specify
missing value.
\item \textbf{\textit{groupcfg}}\textit{=name} - is put into subsections
representnig plugin or {[}group] section and defines if that section
is effective. In old format it selects the group to which all following
lines apply. Only unaffected option is \textbf{\textit{groupcfg}}.
If name is empty (or no groupcfg is used at all) following lines apply
to all users.
\end{itemize}
Subsections of \emph{gridftpd} section specifies plugins which serve
virtual FTP path (similar to mount command of UNIX). Name of subsection
is irrelevant. In old format this section starts with command \textbf{\textit{plugin}}
\textit{path library\_name} and ends with keyword \textbf{\textit{end}}.
Inside subsection following commands are supported

\begin{itemize}
\item \textbf{\textit{plugin}}\textit{=library\_name} - use plugin \textit{library\_name}
to serve virtual path.
\item \textbf{\textit{path}}\textit{=path} - virtual path to serve.

GFS comes with 3 plugins: \textit{filepligin.so, gaclplugin.so} and
\textit{jobplugin.so}.

\begin{itemize}
\item \textit{jobplugin.so} does not require any specific options in case
of old configuration format. It reads the configuration file of the
GM located at the standard place as specified in the section \ref{SubSection: ConfigFile}.
Following options are supported:

\begin{itemize}
\item \textbf{\textit{configfile}}\textit{=path} - defines non-standard
place for GM's configuration file,
\item \textbf{\textit{allownew}}\textit{=yes|no} - specifies if new jobs
can be submitted. Default is \emph{yes}.
\item \textbf{\textit{unixgroup}}/\textbf{\textit{unixvo}}/\textbf{\textit{unixmap}}
- same options like in top-level GFS configuration. If mapping succeeds
obtained local user will be used to run submitted job.
\end{itemize}
\item \textit{filepligin.so} supports following options:

\begin{itemize}
\item \textbf{\textit{mount}}\textit{=path} - defines the place on local
filesystem to which file access operations apply
\item \textbf{\textit{dir}}\textit{=path options} - specifies access rules
for accessing files in \textit{path} (relative to virtual and real
path) and all the files below. \\
\textit{options} is the list of the following keywords:

\begin{itemize}
\item \textbf{\textit{nouser}} - do not use local file system rights, only
use those specifies in this line
\item \textbf{\textit{owner}} - check only file owner access rights
\item \textbf{\textit{group}} - check only group access rights
\item \textbf{\textit{other}} - check only \char`\"{}others\char`\"{} access
rights
\end{itemize}
The options above are exclusive. If none of the above specified usual
Unix access rights are applied.

>>>>

%% LyX 1.4.5.1 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[twoside,english]{article}
\usepackage{pslatex}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{geometry}
\geometry{verbose,a4paper,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2cm}
\usepackage{array}
\usepackage{amsmath}
\usepackage{setspace}
\onehalfspacing
\usepackage{amssymb}
\IfFileExists{url.sty}{\usepackage{url}}
                      {\newcommand{\url}{\texttt}}

@@

A-Rex is one of ARC middleware componets implementing functions of
so called \emph{Computing Element} (CE).

The aim of the   A-Rex is to take care of job pre-
and post-processing. It provides functionality to stage-in files containing
input data and program modules from wide range of sources and transfer
or store output results.

The A-Rex implements Web Service (WS) interface which   
    provides a way to submit and control
computational tasks (jobs) to be executed by the A-Rex and undelying
Local Resource Management System.

@@

Each job gets a directory on the CE called the \emph{session directory}
(SD). Input files are gathered in the SD. The job is supposed to produce
new data files  in the SD as well. The A-Rex does not guarantee
the availability of any other places accessible by the job other than
SD (unless such a place is part of a requested Runtime Environement).
The SD is also the only place which is controlled by the A-Rex.
It is  accessible by the user from outside through HTTP(S)  
protocol. Any file created outside the SD is not controlled by the
A-Rex. Any exchange of data between client and A-Rex (including
also program modules) is done through    HTTP(S) protocol
   . A URL for accessing input/output
files is       obtained through
the WS Local Information Description Interface (LIDI) of A-Rex.

Each job gets an identifier (\textit{jobid}). This is a handle which
identifies the job in the A-Rex and the Information  Interface.

  Jobs are initiated and controlled through   WS interface.
Complete job descriptions   (JD)   are passed to the
A-Rex  through       WS in JSDL \cite{jsdl}
coded description     . Input data files
and job executables are transferred separately through the same interface,
as described in the next section.

@@

One of the most important tasks   of the A-Rex is to
take care of processing input and output data (files) of the job.
Input files are gathered in the SD. There are two ways to put a file
into the SD:

\begin{itemize}
\item Download is initiated by the A-Rex --  This is the case
for files     defined in the JD (with name and source).
      The A-Rex alone is responsible to
make sure that all required files will be available in the SD.\\
The supported protocols for sources  at the moment are: GridFTP,
FTP, HTTP, HTTPS (HTTP over SSLv3) and HTTPg (HTTP over GSI). Also
some nonstandard sources are supported. Those are described below.
The A-Rex fully relies on HED framework \cite{hed} for data transfering
capabilities. Hence actual set of supported protocols depends on installed
Data Management Components of the HED.
\item Upload is initiated by the user directly or through the User Interface
(UI). Because the SD becomes available immediately at the time of
submission of JD, UI can (and should) use that to upload data files
which are not otherwise accessible by the A-Rex.  Examples of
such files  are the main executable of the job,  
the job's input files, etc. These files can (and
should) also be specified in the JD.
\end{itemize}
\textbf{There is no} other reliable way for a job to obtain input
data on the CE based on the A-Rex  . Access to
AFS, NFS, FTP, HTTP and any other remote data transport during execution
of a job is not guaranteed (at least not by A-Rex).

Jobs should store output files in their SD. Like input files,
output files  belong into two groups:

\begin{itemize}
\item Files which are supposed to be moved to a \emph{Storage Element} (SE)
and optionally registered in some \emph{Indexing Service} like the
Globus \emph{Replica Location Service } (RLS)  --
The A-Rex takes care of these files. They have to be specified
in the JD. If the job fails during any stage of processing, no attept
is made to transfer those files to their final destination, unless
the  option \emph{preserve=yes} is specified in their URLs.
\item Files which are supposed to be fetched by the user -- The user has
to use a tool like the UI to obtain these files. They \textbf{must}
also be specified in the JD.
\end{itemize}

@@

From the point of view of the A-Rex a job passes through various
states. Picture \ref{job states diagram} presents a diagram of the
possible states of a job.%
\begin{figure}[tbh]
\begin{centering}
\includegraphics{pic1}
\par\end{centering}

\caption{\label{job states diagram}Job states}
\end{figure}
A user can examine the state of a job by querying the dedicated Local
Information Description Interface of A-Rex  
 using the UI or any other suitable tool

     or through query method of WS interface.

@@

Below is the description of all actions taken by the A-Rex at every
state:

\begin{itemize}
\item \textbf{Accepted} - In this state the job has been submitted to
a CE but is not processed yet. The A-Rex will analyze the JD and
move to the next stage. If the JD can not be processed the job will
be canceled and moved to the state \textbf{Finishing}.
\item \textbf{Preparing} - The input data is being gathered in the SD (stage-in).
The A-Rex is downloading the files specified in the JD and is waiting
for files which are supposed to be uploaded by the UI. If
all files are successfully gathered the job moves to the next state.
If \textbf{any} file can't be downloaded or it takes the UI too long
to upload a file,  the job moves to \textbf{Finishing} state. It
is possible to put a limit on the number of simultaneous \textbf{Preparing}
jobs. If this limit is exceeded, jobs ready to enter the \textbf{Preparing}
state      will stay in the \textbf{Accepted}
state, but prefixed with the PENDING: mark. Exceptions are jobs which
have no files to be downloaded. These are processed out of limits.
\item \textbf{Submitting} -      The job is being
passed for execution to the \emph{Local Resource Management System}
(LRMS).         The correspoding backends
for many LRMSs are provided with the default installation.  
     If the local job submission is successful
the job moves to the \textbf{InLRMS} state. Otherwise it moves
to \textbf{Finishing}. It is possible to limit the agregate number
of jobs in \textbf{Submitting} and  \textbf{InLRMS} states.
\item \textbf{InLRMS} - The job is queued or being executed in the LRMS.
The A-Rex takes no actions except waiting until the job finishes.
\item \textbf{Cancelling} - Necessary action to cancel the job in the LRMS
is being taken.
\item \textbf{Finishing} - The output data is being processed (stage-out).
Specified data files are moved to the specified SEs and are optionally
registered at  an Indexing Service. The user can download data files
from the SD by using the UI or  other adequate tool. All the files
not specified as output files are removed from the SD at very beginning
of this state. It is possible to limit number of simultaneous jobs
in this state.
\item \textbf{Finished} - No more processing is performed by the A-Rex.
The user can continue to download data files from the SD. The SD is
kept available for some time (default is 1 week). After that the
job is moved to the state \textbf{Deleted}. The 'deletion' time can
be obtained by querying the Informatrion Interface of the A-Rex
      . If a job was moved to \textbf{Finished}
because of failure, it may be restarted on request of a client. When
restarted, a job is moved to the state previous to the one in which
it failed and is assigned mark PENDING. This is needed in order to
not break the configuration limits. Exception is a job failed in \textbf{InLRMS}
state and lacking input files specified in JD. Such a job is treated
like failed in \textbf{Preparing} state.
\item \textbf{Deleted} - The job is moved to this state if the user have
not requested job to be cleaned before the SD's lifetime expires.
Only minimal subset of information about such job is kept. The SD
is not available anymore.
\end{itemize}
In  case of a failure special processing is applied to output
files. All specified output files are treated as \textbf{downloadable
by the user}. No files will be moved to their destination SE.

@@

In full installation the A-Rex and it's components support the
following data transfer protocols and corresponding URLs: \emph{ftp,
gsiftp, http, httpg, https, se, rc} and \emph{rls.} For more information
please see {}``The Hosting Environment of the Advanced Resource Connector
middleware      
  '' document \cite{hed}.

@@

\subsection{Internal Files of the A-Rex }

For each local UNIX user listed in the A-Rex configuration - including
a generic one which covers all local user identities - a \textit{control
directory} exists. In this directory the A-Rex stores information
about jobs belonging to that user. Multiple users can share the same
\textit{control directory}. To make it easier to recover in  case
of failure, the A-Rex stores most information in files rather than
in memory. All files belonging to the same job have names starting
with \textbf{job.ID.,} where ID is the job identifier.

@@

\begin{itemize}
\item \textit{job.ID.description} - contains the  description of the
job (JD).
\item \textit{job.ID.local} - information about the job used by the A-Rex.
It consists of lines of format \textit{{}``name = value''} . Not
all of them are always available. The following names are defined:

@@

\item \textit{clientsoftware} - version of software used to submit the job
\item \textit{sessiondir} - the job's SD  
\item \textit{failedstate} - state in which job failed (available only
if it is possible to restart the job)
\item \textit{jobreport} - URL of a user requested \emph{logger service}.
The GM will also send job records to this service in addition to the
default logger service configured in the configuration  
        (This feature does
not work in current implementation).
\end{itemize}
\end{itemize}
This file is filled partially during job submission and fully when
the job moves from the \textbf{Accepted} to the \textbf{Preparing}
state.

@@

\item \textit{job.ID.proxy} -delegated X509 credentials .
\item \textit{job.ID.proxy.tmp} - temporary X509 credentials  with
different UNIX ownership used by processes run with effective \emph{user
id} different form job owner's \emph{id}.
\end{itemize}
There are other files with names like job.ID.{*} which are created
and used by different parts of the A-Rex. Their presence in the
\textit{control directory} can not be guaranteed and can change depending
on changes in the A-Rex code.

\subsection{ }

    \textit{ }  \textit{}

 \textit{}

 \textit{}

 \textit{}

@@

A-REX Web Service Interface provides means to submit a description
of a computational job to a computing resource, to stage-in additional
data, to monitor and control processing of jobs, and obtain data produced
during the execution of a job. The WS Interface is built and deployed
inside Hosting Envirnoment Daemon (HED) infrastructure\cite{hed}.

@@

The job submission and control interface is based on document produced
by the OGF OGSA Basic Execution Services (BES) Working Group\cite{bes}.
The exchange of SOAP messages is performed over the HTTP(S) protocol.
The BES interface is represented by two port-types - BES-Management
and BES-Factory. The former is made to control the A-REX service itself
and thus defines operations to start and stop the functionality of
the BES service. The A-Rex does not implement remote control of
service functionality. Hence the BES-Management port-type is not functional.
The BES-Factory port-type provides operations to submit new jobs (to
create an activity in terms of BES) and to monitor its state. It also
has an ability to provide information about the service. A-Rex fully
implements the functionality of this port-type.

For job descriptions A-Rex accepts the Job Submission Description
Language (JSDL)\cite{jsdl} documents as defined by the OGF Job Submission
Description Language Working Group. Supported elements and extensions
are described below.

@@

A-R introduces a new operation in addition to those provided by
BES. It does that by defining its own port-type  with the single
operation \emph{ChangeActivityStatus} (see Appendix D). This operation
provides a way to request simple transfers between states of jobs
  and corresponding actions.

@@

Since BES allows implementators to extend their initial activity states
with additional sub-states, A-Rex defines a set of sub-states of
activity processing in addition to those defined by the BES, as listed
in Table \ref{tab:Job-states-definitions}.

@@

       The A-Rex
also supportd  the Delegation Interface (see Appendix E). This
is a common purpose interface to be used by ARC services which accepts
delegated credentials from clients. The Delegation Interface implements
two operations: initiatialization of credentials delegation (DelegateCredentialsInit)
and update/renewal of credentials (UpdateCredentials).

@@

The A-REX implements the Local Information Description Interface (LIDI)
interface  common for all ARC services. This interface is based
on OASIS Web Services Resource Properties specification\cite{wsrf-rp}.
Information about resources and maintained activities/jobs are represented
in a \emph{WS-Resource Properties} informational XML document. The
document type is defined in the A-Rex WSDL as a \emph{ResourceInformationDocumentType}.
It contains the following elements/resources:

@@

\emph{grid-manager {[}-h] {[}-C level] {[}-d level] {[}-c path] {[}-F]
{[}-U uid{[}:gid]] {[}-L path] {[}-P path]}\\
\hspace*{0.5cm}\emph{-h} - short help,\\
\hspace*{0.5cm}\emph{-d} - debug level,\\
\hspace*{0.5cm}\emph{-L} - name log file (overwrites value in configuration
file),\\
\hspace*{0.5cm}\emph{-P} - name for file containing process id (overwrites
value in configuration file),\\
\hspace*{0.5cm}\emph{-U} - user and gropu id to use for running daemon,\\
\hspace*{0.5cm}\emph{-F} - do not make process daemon,\\
\hspace*{0.5cm}\emph{-c} - name od configuration file,\\
\hspace*{0.5cm}\emph{-C} - remove old information before starting:
1- remove finished jobs, 2 - remove active jobs too, 3- also remove
everything that looks like junk.

\subsection{    }

        \textit{}
 \textit{}

\begin{itemize}
\item \textbf{\textit{}}\textit{}

\item \textbf{\textit{}}\textit{}

\item \textbf{\textit{}}\textit{}

\item \textbf{\textit{}}\textit{}

\item \textbf{\textit{}}\textit{}     \emph{}

\item \textbf{\textit{}}\textit{}      
\textit{}

  \textit{}

\item \textbf{\textit{}}\textit{ }

 \textit{}       
        \textit{}

\begin{itemize}
\item \textbf{\textit{}} \textit{}

\item \textbf{\textit{}} \textit{}    
         \textit{}

\textit{}

\item \textbf{\textit{}} \textit{  }  
      \textit{}    
        \textit{}
\textit{}       
   \textit{}     
       \textit{}

\item \textbf{\textit{}} \textit{ } \textit{}
\textit{}     \emph{}  
    \emph{}    
 \emph{}          
     \emph{}

\begin{itemize}
\item      
\item       
\end{itemize}
\end{itemize}
\item \textbf{\textit{}}\textit{ }    \textbf{\textit{}}
       \textit{}
\item \textbf{\textit{}}\textit{ }

\textit{}       \textbf{\textit{
}}      \textbf{\textit{}}\textbf{\textit{}}

 \textit{}    \textit{}
 \textit{}

\item \textbf{\textit{}}\textit{}

      \textbf{\textit{}}

\end{itemize}
  \emph{}

          \textbf{\textit{}}
\textit{ }     \textbf{\textit{}}

\begin{itemize}
\item \textbf{\textit{}}\textit{}    \textit{}

\item \textbf{\textit{}}\textit{}

     \textit{ } 
\textit{}

\begin{itemize}
\item \textit{}

\begin{itemize}
\item \textbf{\textit{}}\textit{}

\item \textbf{\textit{}}\textit{}     
     \emph{}
\item \textbf{\textit{}}\textbf{\textit{}}\textbf{\textit{}}

\end{itemize}
\item \textit{}

\begin{itemize}
\item \textbf{\textit{}}\textit{}

\item \textbf{\textit{}}\textit{ }    
    \textit{}     
      \textit{}

\begin{itemize}
\item \textbf{\textit{}}

\item \textbf{\textit{}}       
\item \textbf{\textit{}}      
\item \textbf{\textit{}}      
\end{itemize}

\begin{itemize}
\item \textbf{\textit{}}    
\item \textbf{\textit{}}    
\item \textbf{\textit{}}        
\item \textbf{\textit{}}

\item \textbf{\textit{}}       
\item \textbf{\textit{}}       
\item \textbf{\textit{}} \textit{ }
          \textit{} 
    \textit{}

   \textit{}  \textit{}

\item \textbf{\textit{}} \textit{ }

\end{itemize}
\end{itemize}
\item \textit{}

      \textbf{\textit{}}

   \textbf{\textit{}}\textit{}  
 \textbf{\textit{}}\textit{}

\begin{lyxlist}{00.00.0000}
\item [{\emph{}}]      
\item [{\emph{}}]      
\item [{\emph{}}]       
\item [{\emph{}}]     
\item [{\emph{}}]     
\item [{\emph{}}]        
\end{lyxlist}
     \emph{}

\end{itemize}
\end{itemize}

\emph{         
          }\emph{}
  \emph{}   \emph{}    
    \emph{}    
       \emph{}
         \emph{}   
  \emph{}     \emph{}
   \emph{}     
 \emph{}        
   \emph{}
